Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-09-04 prop_letter.tex
prop_letter.tex \section{Kochen-Specker} Add the suggestions from Gerardo...

\noindent The observables are constructed using the Pauli matrices
$\left\{ \mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ whose
eigenvalues are all either 1 or -1. They are arranged such that in
each row and column, \emph{except the column $j=2$}, every observable
is the product of the other two. In the $j=2$ column, we have instead
that
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$. Now
assume a QIVPM ${\bar{\mu}}$; the expectation values of the
observables in each row relative to this QIVPM are fully determinate
and must lie in either the interval $[1,1]$ or the interval $[-1,-1]$
depending on which eigenvalue is the one whose associated projector is
certain. Since the product of any two observables in a row is equal to
the third, there must be an even number of occurrences of the interval
$[-1,-1]$ in each row and hence in the entire table. However, looking
at the expectation values of the observables in each column, there
must be an even number of occurrences of the interval $[-1,-1]$ in the
first two columns and an odd number in the $j=2$ column and hence in
the entire table. The contradiction implies the non-existence of the
assumed QIVPM.

@@

Our framework allows us to generalize the above theorem to state that
for small enough $\delta$, it is impossible to have
$\delta$-deterministic QIVPMs. Every QIVPM must map some events to
truly uncertain intervals, not just ``almost definite intervals.'' The
significance of this observation is that the uncertainty in the
intervals is, in contrast to the classical case, \emph{a fundamental
  part of quantum theory}. There are in fact two sources of
uncertainty: one due to the finite measurement resources and one due
to complementarity, which requires every quantum framework to be truly
probabilistic.

@@

When $\delta \ge \frac{1}{3}$, the situation abruptly
changes. Assuming a Hilbert space $\Hilb$ of dimension $3$, then for
any projector $P$ that projects onto an $n$-dimensional subspace of
$\Hilb$, it is straightforward to check that
$\bar{\mu}\left(P\right)=\left[n/3,n/3\right]$ is a valid
$\frac{1}{3}$-deterministic QIVPM. In other words, when the uncertainty
in measurements becomes so large, the Kochen-Specker theorem no longer
holds and it becomes possible to map every observable to some (quite
inaccurate) probability interval.

>>>>

\noindent The observables are constructed using the Pauli matrices
$\left\{ \mathbb{1},\sigma_{x},\sigma_{y},\sigma_{z}\right\}$ whose
eigenvalues are all either 1 or $-1$. They are arranged such that in
each row and column, \emph{except the column $j=2$}, every observable
is the product of the other two. In the $j=2$ column, we have instead
that
$\left(\sigma_{z}\otimes\sigma_{z}\right)\left(\sigma_{x}\otimes\sigma_{x}\right)=-\sigma_{y}\otimes\sigma_{y}$. Now
assume a 0-deterministic QIVPM ${\bar{\mu}}$; the expectation values of the
observables in each row relative to this QIVPM are fully determinate
and must lie in either the interval $[1,1]$ or the interval $[-1,-1]$
depending on which eigenvalue is the one whose associated projector is
certain. Since the product of any two observables in a row is equal to
the third, there must be an even number of occurrences of the interval
$[-1,-1]$ in each row and hence in the entire table. However, looking
at the expectation values of the observables in each column, there
must be an even number of occurrences of the interval $[-1,-1]$ in the
first two columns and an odd number in the $j=2$ column and hence in
the entire table. The contradiction implies the non-existence of the
assumed QIVPM.

@@

Our framework allows us to generalize the above theorem to state that
for small enough $\delta$, it is impossible to have
$\delta$-deterministic QIVPMs. We therefore prove a stronger statement of 
contextuality that includes the effects of finite-precision.
Every QIVPM must map some events to
truly uncertain intervals, not just ``almost definite intervals.'' The
significance of this observation is that the uncertainty in the
intervals is, in contrast to the classical case, \emph{a fundamental
  part of quantum theory}. There are in fact two sources of
uncertainty: one due to the finite measurement resources and one due
to complementarity, which requires every quantum framework to be truly
probabilistic.

@@

When $\delta \ge \frac{1}{3}$, the situation abruptly
changes. Assuming a Hilbert space $\Hilb$ of dimension $3$, then for
any projector $P$ that projects onto an $n$-dimensional subspace of
$\Hilb$, it is straightforward to check that
$\bar{\mu}\left(P\right)=\left[n/3,n/3\right]$ is a valid
$\frac{1}{3}$-deterministic QIVPM. In other words, when the uncertainty
in measurements becomes so large, the Kochen-Specker theorem no longer
holds and it becomes possible to map every observable to some (quite
inaccurate) probability interval and hence the theory becomes non-contextual.
